\documentclass{article}
\usepackage[final]{neurips_2021}
\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{xcolor}         % colors


\title{Project Report for UMC 203: \\
Flow Based Generative Models
(working version)}

\author{%
  Akshaya Garepally \\
  25398 \\
  \texttt{gakshaya@iisc.ac.in}
    \And
  Archana Yadav \\
  25315 \\
  \texttt{archanay@iisc.ac.in.in}
    \And
  Himanshi Sarode \\
  25406\\
  \texttt{himanshis@iisc.ac.in}
  \And
  Sri Varshinee S \\
  24484 \\
  \texttt{varshinees@iisc.ac.in}
  \And
  Swetha Nomula \\
  25391 \\
  \texttt{swethan@iisc.ac.in.in}
}

\begin{document}

\maketitle


\section{Reproduction of Experiments}


\section*{References}


\small

[1] Dinh, L.\ Krueger, D.\ \& Bengio, Y.\ (2014) NICE: Non-linear Independent Components Estimation. {\it arXiv preprint arXiv:1410.8516}.

[2] Dinh, L.\ Sohl-Dickstein, J.\ \& Bengio, S.\ (2016) Density estimation using Real NVP. {\it arXiv preprint arXiv:1605.08803}.

[3] Kingma, D.P.\ \& Dhariwal, P.\ (2018) Glow: Generative Flow with Invertible 1x1 Convolutions. {\it arXiv preprint arXiv:1807.03039}.

[4] Wang, L.\ Li, S.\ Yang, F.\ Wang, J.\ Zhang, Z.\ Liu, Y.\ Wang, Y.\ \& Yang, J.\ (2025) Not All Parameters Matter: Masking Diffusion Models for Enhancing Generation Ability. {\it arXiv preprint arXiv:2505.03097}.

[5] Scott, D.W.\ (1992) {\it Multivariate Density Estimation: Theory, Practice, and Visualization}. New York: John Wiley \& Sons.

[6] Papamakarios, G.\ Nalisnick, E.\ Rezende, D.J.\ Mohamed, S.\ \& Lakshminarayanan, B.\ (2021) Normalizing Flows for Probabilistic Modeling and Inference. {\it Journal of Machine Learning Research} 22(57), pp.\ 1--64.

[7] McInnes, L.\ Healy, J.\ \& Melville, J.\ (2018) UMAP: Uniform Manifold Approximation and Projection for Dimension Reduction. {\it arXiv preprint arXiv:1802.03426}.

[8] van der Maaten, L.\ \& Hinton, G.\ (2008) Visualizing Data using t-SNE. {\it Journal of Machine Learning Research} 9, pp.\ 2579--2605.

[9] Irwin, R.\ Tibo, A.\ Janet, J.-P.\ \& Olsson, S.\ (2024) Efficient 3D Molecular Generation with Flow Matching and Scale Optimal Transport. {\it arXiv preprint arXiv:2406.07266}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\end{document}
